---
title: "A1_Jewitt_20096665"
author: "Ally Jewitt"
date: "10/01/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
```

## Set theme

```{r}
source("./ThemePub.R")
```

## Run R scripts

This script creates 4 vectors for Organism, Limb_width, Limb_Length, and Observer, and arranges them into a data.frame.

```{r}
MyData<-data.frame(Organism=sample(c("Ailuropoda melanoleuca", "Panthera tigris", "Elephas maximus", "Gorilla beringei", "Pongo pygmaeus"), 100, replace=T), Limb_width=rnorm(100, 40, 5), Limb_length=rnorm(100, 70, 5), Observer=sample(c("Dr. Jane Goodall", "Dian Fossey", "Dr. Biruté Galdikas"), 100, replace=T))
```

This script adds a Volume column to the measurements.csv file using the equation for the volume of a cylinder (πr\^2h).

```{r}
MyData<-MyData %>% 
  mutate(Volume=pi*((MyData$Limb_width/2)^2)*MyData$Limb_length)
```

## Link to GitHub

[GitHub repository](https://github.com/allyjewitt/BIOL432_Assignment1)

## 1. Load the revised measurements.csv

```{r}
head(MyData)
```

## 2. Use dplyr to sort the data by Organism, then by Observer, then by Volume

```{r}
MyData <- MyData %>% 
  arrange(Organism, Observer, Volume)
head(MyData)
```

## 3. Create a table showing the average volume for each species

```{r}
AvgVolume<-aggregate(MyData$Volume, list(MyData$Organism), mean)
AvgVolume %>% 
  rename(
    Mean = x,
    Organism = Group.1
  )
```

## 4. Create a table showing the number of observations (i.e. rows) for each combination of species and observer.

```{r}
NumObservations<- MyData %>% group_by(Organism, Observer) %>% 
  summarize(NumObservations = n())
NumObservations 
```

## 5. Create a single box plot (or similar) to compare key features of the distributions of volumes for each species.

```{r}
ylab <- expression(Volume ~ cm^3)
ggplot(aes(x=Organism, y=Volume), data=MyData) + geom_boxplot() + ylab(ylab) + theme_pub()
```

Figure 1. Box plot comparing the key features of the distribution volumes of each species. The boxes represent the interquartile range, and the horizontal line represents the median value. Upper and lower tails represent the maximum and minimum values. The dots represent any outliers. Each species has a similar median value for limb volume.

## 6. Create a multi-panel plot showing frequency histograms for each species.

```{r}
xlab <- expression(Volume ~ cm^3)
ggplot(MyData, aes(Volume)) + geom_histogram(bins=7) + facet_wrap(.~Organism) + xlab(xlab) + ylab("Count") + theme_pub()
```

Figure 2. Multi-panel plot showing the frequency histograms for the limb volume for each species, *Ailuropoda melanoleuca*, *Elephas maximus*, *Gorilla beringei*, *Panthera tigris*, and *Pongo pygamaeus*. Each histogram shows a similar distribution for limb volume with the highest frequency at the mean value.
